---
title: "Final Presentation"
author: "Connor Wang"
subtitle: "5/8/25"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    self-contained: false
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Project 1 - Data Visualization

::: panel-tabset
### Overview

-   Two TidyTuesday datasets

-   Data on Bob's Burgers Dialogue from various episodes.

-   Data on how Horror Movies performed in the box office/how much the public enjoyed them.

### Bob's Burgers

-   American Sitcom with 15 seasons

-   Follows Bob Belcher and his family, who own the restaurant Bob's Burgers

    ```{r echo = FALSE, results = "hide", message=FALSE, warning=FALSE}
    library(tidyverse)

    episode_metrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-19/episode_metrics.csv')

    episode_metrics
    ```

    ```{r echo=FALSE}
    head(episode_metrics)
    ```

-   Explored a possible relationship between the amount of lines that ended with questions versus exclamations

    ```{r message=FALSE, warning=FALSE}
    #| output-location: column
    ggplot(episode_metrics, aes(x = question_ratio, y = exclamation_ratio)) +
      geom_point(color = "blue") + 
      labs(
        title = "Question Ratio vs. Exclamation Ratio in Bob's Burgers episodes",
        x = "Ratio of lines with questions",
        y = "Ratio of lines with exclamations"
      )
    ```

-   No clear trends from scatterplot, more questions does not mean more excitement in an epsiode.

### Horror Movies

-   Popular genre of movies meant to evoke feelings of fear and terror

-   Since 2017, around 40+ horror movies are released each year

    ```{r results = "hide", echo=FALSE, message=FALSE, warning=FALSE}
    library(tidyverse)
    horror_movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-11-01/horror_movies.csv')

    horror_movies
    ```

    ```{r echo=FALSE}
    head(horror_movies)
    ```

-   Explored relationship between a movie's budget and revenue

    ```{r message=FALSE, warning=FALSE}
    #| output-location: column

    ggplot(horror_movies, aes(x = budget, y = revenue)) +
      geom_point(color = "red") +
      labs(
        title = "Horror Movie Budget versus Revenue",
        x = "Budget",
        y = "Revenue"
      )
    ```

-   No clear trends from scatterplot, spending more money does not necessarily mean you will make more money.
:::

## Project 2 - Netflix Data Analysis

```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-04-20/netflix_titles.csv')

netflix_titles
```

-   Netflix is a popular streaming service consisting of TV shows, documentaries, movies, and more.

```{r echo=FALSE, warning=FALSE, message=FALSE}
head(netflix_titles)
```

-   Created three questions of interest regarding the information from the dataset.

## Has movie duration changed over time?

-   Data from the mid 1900s to 2022.

-   Films from multiple countries and genres on Netflix.

-   Manipulated data to include only movies that have a runtime of a hour or longer.

    ```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
    library(tidyverse)
    library(ggplot2)
    netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-04-20/netflix_titles.csv')

    netflix_titles
    ```

    ```{r, warning=FALSE, message=FALSE}
    movies_only <- netflix_titles %>%
      filter(type == "Movie") %>%
      mutate(movie_duration = as.numeric(str_extract(duration, "\\d+"))) %>%
      filter(!is.na(duration) & movie_duration >= 60)%>%
      arrange(release_year)%>%
      select(type, release_year, duration, movie_duration)

    head(movies_only)
    ```

    ```{r, warning=FALSE, message=FALSE}
    ggplot(movies_only, aes(x = release_year, y = movie_duration)) + 
      geom_point(color = "lightsalmon3") +
      labs(
        title = "Has Movie duration changed over time?",
        x = "Year Released",
        y = "Duration of Movie in Minutes(starting at 60)"
      )
    ```

-   Found no trend, movie durations have stayed in a similar range over the years.

## Movies and Shows about the future

-   Dataset included a brief description about each piece of media.

-   Analyzed the number of films, TV shows, documentaries, etc. that were about the future categorized by country.

    ```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
    library(tidyverse)
    library(ggplot2)
    netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-04-20/netflix_titles.csv')

    netflix_titles
    ```

    ```{r, message=FALSE, warning=FALSE}
    content_about_future <- netflix_titles %>%
      filter(str_detect(description, "(?i)future")) %>%
      filter(!is.na(country)) %>%
      group_by(country) %>%
      summarise(count = sum(n())) %>%
      arrange(desc(count)) %>%
      slice_head(n = 10)

    head(content_about_future)
    ```

    \`\`\`{ggplot(content_about_future, aes(x = country, y = count)) +} geom_col(fill = "plum3") + labs( title = "Countries with TV Shows and Movies about the 'future'", x = "Country", y = "Number of TV Shows/Movies" ) + theme(axis.text.x = element_text(size = 7))

    \`\`\`

-   United States had the most, but likely because most of the media was from the US to begin with.

-   Most countries are similar, no country is overly creative about the future.

## Words preceding 'of'

-   Dataset included titles of many different works of media.

-   Analyzed which words most commonly came before 'of' in a title.

-   Chose 'of' because of how common it is used - ex: story of, legend of, etc.

    ```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
    library(tidyverse)
    library(ggplot2)
    netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-04-20/netflix_titles.csv')

    netflix_titles
    ```

    ```{r, message=FALSE, warning=FALSE}
    words <- netflix_titles %>%
      mutate(of = str_extract(str_to_lower(title), "\\b\\w+(?= of)")) %>%
      filter(!is.na(of)) %>%
      group_by(of) %>%
      summarise(n = n()) %>%
      arrange(desc(n)) %>%
      slice_head(n = 10)
    head(words)
    ```

    ```{r, message=FALSE, warning=FALSE}
    ggplot(words, aes(x = of, y = n)) +
      geom_col(aes(fill = of)) + 
      labs(
        title = "10 Most common words that come before 'of' in TV show/movie titles",
        x = "Word before 'of'",
        y = "Number of titles that it appears before 'of' in"
      )
    ```

-   Most common words were (as predicted) legend, story.

-   Similar movie title formats are recycled.

# Thank You! {background-color="lightblue"}
