---
title: "Netflix Data Analysis"
description: |
  Full Text Analysis of Netflix Title Data
author: Connor Wang
date: March 2, 2025
format: html
---

```{r, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```

### Dataset

The Dataset that I have chosen to explore is about Netflix shows. Netflix is a streaming service which includes various documentaries, TV shows, movies, and more. The data consists of TV shows and movies that were available on Netflix from 2019 to 2021. The data was collected through Flixable, a third-party Netflix search engine. The data includes the type of picture it is, such as movie or TV show, the title, the director, the cast, the country it originates from, the date it was added to netflix, the year it was released, the ratings(TV-MA, PG-13, etc.), the duration, what categories it is listed in, and the description of the picture. Through this information I have created various questions of interest that come with accompanying visuals in order to share my findings.

```{r, results='hide', warning=FALSE, message=FALSE}
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-04-20/netflix_titles.csv')

netflix_titles
```

### Question of Interest 1

The first question of interest that I came up with for this dataset is how has movie duration changed over time? Since the data about what is streaming on Netflix included the duration of movies in minutes as well as the year they were released, I decided it would be interesting to note how this has changed over time. Going into it I assumed that movies have gotten longer over time. Movies that come out these days seem to always be two plus hours, often being two and a half hours or even three on some of the extreme cases. Although not all movies are included on Netflix, it would still give a good idea of how movies have trended in terms of duration over time. I found the opposite to be true. Based on the linear regression model I created below there is a clear downwards trend over time. Movies have in fact gotten shorter in duration throughout the years and. Save for some outliers, the general trend is the same with newer movies becoming shorter and shorter over time.

```{r, warning=FALSE, message=FALSE}
movies_only <- netflix_titles %>%
  filter(type == "Movie") %>%
  mutate(movie_duration = as.numeric(str_extract(duration, "\\d+"))) %>%
  filter(!is.na(duration))

ggplot(movies_only, aes(x = release_year, y = movie_duration)) + 
  geom_point(color = "lightsalmon3") +
  geom_smooth(method = "lm", color = "steelblue3", se = FALSE) + 
  labs(
    title = "Has Movie duration increased over time?",
    x = "Year Released",
    y = "Duration of Movie in Minutes"
  )
```

Previously mentioned in the description of the question of interest, this plot depicts the release year of a movie on the x-axis versus the duration of the movie in minutes on the y-axis. This creates a comparison between movie duration over time. You see a negatively-sloped linear regression line, indicating that movie duration time has decreased in general over the years. Furthermore, more movies in general have come out as time has progressed which could be a part of it. There are more shorter movies out there as they are likely easier to produce than longer ones.

### Question of Interest 2

The second question of interest that I came up with was which countries have the most TV Shows and Movies that deal with the future. I wondered if there would be a connection to depictions of the future with countries that tend to be seen as more technology advanced or first world. I explored the descriptions that included the word future in them and used this to determine if the picture had anything to do with the future. What I found generally aligned with what I thought as the United States had the highest amount of TV shows and movies dealing with 'future' ideas. This is also likely why I thought about this because I am exposed to so many shows and movies that align with this, having grown up watching Netflix in the states. The bar chart shows a large discrepancy from the United States to the next country, India. This uses only the top 10 countries that have future in descriptions.

```{r}
content_about_future <- netflix_titles %>%
  filter(str_detect(description, "(?i)future")) %>%
  filter(!is.na(country)) %>%
  group_by(country) %>%
  summarise(count = sum(n())) %>%
  arrange(desc(count)) %>%
  slice_head(n = 10)

ggplot(content_about_future, aes(x = country, y = count)) +
  geom_col(fill = "plum3") + 
  labs(
    title = "Countries with TV Shows and Movies about the 'future'",
    x = "Country",
    y = "Number of TV Shows/Movies"
  ) +
  theme(axis.text.x = element_text(size = 7))
```

Previously mentioned in the description of the question of interest, this plot shows the top 10 countries that have TV Shows or Movies on Netflix with future in their description. The y-axis shows how many TV Shows or Movies follow this criteria. Outside of the United States, there are not very many TV Shows or Movies that are about future scenarios. The US is the clear leader by about 20 projects. The next is India with about 8 or 9. Most have below 5.

### Question of Interest 3

The third question of interest that I came up with was which words come before the word 'of' the most in movie titles. I thought about this because there seem to be many shows and movies that use of, such as Pirates of the Caribbean, Wizard of Oz, and many more. I found this would be an interesting thing to see. I did not have much prediction going in, but after seeing the data it makes sense. Words such as story and legend tend to come before of. Many title creators likely see it as a good way of making a fantastical-sounding title, such as the story of \_\_\_ or the legend of \_\_\_. I used the title column in the data to create a lookaround that checked for ' of' following the word. This is how I got the 10 most used words before 'of', finding an interesting trend that I do not think many people would tend to think of.

```{r}
words <- netflix_titles %>%
  mutate(of = str_extract(str_to_lower(title), "\\b\\w+(?= of)")) %>%
  filter(!is.na(of)) %>%
  group_by(of) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice_head(n = 10)

ggplot(words, aes(x = of, y = n)) +
  geom_col(fill = "royalblue3") + 
  labs(
    title = "10 Most common words that precede 'of' in TV show/movie titles",
    x = "Word before 'of'",
    y = "Number of titles that it appears before 'of' in"
  )
```

Previously mentioned in the question of interest, this plot shows the top 10 words that come before 'of' in the titles of TV shows or movies. The words legend and story are both tied for first place the most amount of times it comes before the word 'of' in a title. The 7 other words also appear in front of 'of' a considerable amount of times, all of which having at least 6 projects in which they are used in the title prior to the use of 'of'.

### References

Bansal, Shivam. “Netflix Movies and TV Shows.” Kaggle, 27 Sept. 2021, www.kaggle.com/datasets/shivamb/netflix-shows.

Rfordatascience. “Tidytuesday/Data/2021/2021-04-20/Readme.Md at Main · Rfordatascience/Tidytuesday.” GitHub, github.com/rfordatascience/tidytuesday/blob/main/data/2021/2021-04-20/readme.md. Accessed 2 Mar. 2025.
