---
title: "Arrests and Warnings categorized by race across the states of Arizona, Maryland, and Colorado"
description: |
  SQL queries of the data from Stanford Open Policing Project
author: Connor Wang
date: April 28, 2025
format: html
---

```{r warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: "R Packages"
library(ggplot2) # R package
```

```{r warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: "Establish Database Connection"
con_traffic <- DBI::dbConnect( # connection to database
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)
```

## Introduction

The Stanford Open Policing Project(2020) is a team of researchers and journalists at Stanford University. With the goal of improving investigation and interaction between the police and the public, this team of researchers, including Pomona College's own Amy Shoemaker (2020), gathered records from millions of traffic stops. They then analyze and release records from these stops. When a traffic stop is made, a cop can choose to arrest an individual, write them a citation, or give them a warning. A warning is the most minor, while a citation tends to mean a fine and other possible punishments, with an arrest being the worst as there is more reason to believe that a true crime has been committed. I thought this was interesting and important to look at to see if there are any trends based on the severity of the punishment given to an individual and their race. Given the diversity of the United States of America, I chose to use 3 states that represented different areas of the country - the west, through Arizona, the midwest through Colorado, and the east coast, through Maryland. Although they have different populations, with Arizona over a million people more than the second highest population, Maryland, I thought it was best to get a good representation of different regions of the country. I filtered and used the data to create rates of arrests, warnings, and citations given across races. I then created bar graphs to display differences across states based on this information.

## Arrests

```{sql}
#| connection: con_traffic
#| output.var: "arrests_table"
#| code-fold: true
#| code-summary: "Arrests Table"

-- Maryland arrests
SELECT 'Maryland' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(arrest_made) AS num_arrests,
  SUM(arrest_made) * 1.0 / COUNT(*) AS rate_of_arrests -- calculate rate of arrests
FROM md_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 

UNION

-- Arizona arrests
SELECT 'Arizona' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(arrest_made) AS num_arrests,
  SUM(arrest_made) * 1.0 / COUNT(*) AS rate_of_arrests -- calculate rate of arrests
FROM az_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 

UNION

-- Colorado arrests
SELECT 'Colorado' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(arrest_made) AS num_arrests,
  SUM(arrest_made) * 1.0 / COUNT(*) AS rate_of_arrests -- calculate rate of arrests
FROM co_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
head(arrests_table)
```

The SQL query creates a table with information regarding the rates of arrests made from traffic stops in Maryland, Colorado, and Arizona, categorized by race. The information is filtered to remove data that did not have a race associated with it, such as 'other' and 'unknown'. It is compiled into one table, represented by the tibble above. The main columns we are looking at here are state, subject_race, and rate_of_arrests. This is what we use to create our visualization of the data.

```{r warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: "Plot of arrests"
#| fig-alt: "The bar graph represents the rate of arrests made across the races Black, Hispanic, Asian/Pacific Islander, and White in the states Maryland, Arizona, and Colorado. Arizona issues the highest rate of warnings while Colorado issues the least for Asian/Pacific Islander, Hispanic, and White. Maryland issues the least for Black." 
ggplot(arrests_table, aes(x = subject_race, y = rate_of_arrests, fill = state)) +
  geom_col(position = "dodge") + # bar plot with rate of arrests and race
  labs( # plot labels
    title = "Rate of arrests by race across the states \n
    Arizona, Colorado, and Maryland",
    x = "Race",
    y = "Rate of Arrests"
  )
```

The bar graph represents the rate of arrests made by race, with each race having three bars representative of the different states. Arizona has the highest rate of arrests made for every race, with Black individuals having the highest arrest rate. Maryland and Colorado had similar rates of arrests across all races, both having Asians/Pacific Islanders with the lowest rate of arrests and Hispanics having the highest rate of arrests out of these recorded traffic stops.

## Citations

```{sql}
#| connection: con_traffic
#| output.var: "citations_table"
#| code-fold: true
#| code-summary: "Citations Table"

-- Maryland citations
SELECT 'Maryland' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(citation_issued) AS num_citations,
  SUM(citation_issued) * 1.0 / COUNT(*) AS rate_of_citation -- calculate citation rates
FROM md_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 

UNION

-- Arizona citations
SELECT 'Arizona' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(citation_issued) AS num_citations,
  SUM(citation_issued) * 1.0 / COUNT(*) AS rate_of_citation -- calculate citation rates
FROM az_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 

UNION

-- Colorado Citations
SELECT 'Colorado' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(citation_issued) AS num_citations,
  SUM(citation_issued) * 1.0 / COUNT(*) AS rate_of_citation -- calculate citation rates
FROM co_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
head(citations_table)
```

The SQL query creates a table with information regarding the rate of citations given from traffic stops in Maryland, Colorado, and Arizona, categorized by race. The data is filtered to remove races in the dataset labelled as 'other' or 'unknown'. It is compiled into one table, represented by the tibble above. The data includes subject_race, state, and rate_of_citation which are used to create the visualization.

```{r warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: "Plot of citations"
#| fig-alt: "The bar graph represents the rate of citations given across the races Black, Hispanic, Asian/Pacific Islander, and White in the states Maryland, Arizona, and Colorado. Arizona issues the highest rate of citations while Colorado issues the least across all races." 
ggplot(citations_table, aes(x = subject_race, y = rate_of_citation, fill = state)) +
  geom_col(position = "dodge") + # bar plot using rate of citation and subject race
  labs( # graph labels
    title = "Rate of citations given by race across the states \n 
    Maryland, Arizona, and Colorado",
    x = "Race",
    y = "Rate of Citations Issued"
  )
```

Similar to the arrests data, Arizona has the highest rate of citations issued across all races. This time, Asians/ Pacific Islanders are recorded to have the highest rate of citations issued, while White individuals have the lowest. Maryland is similar, with White individuals having the lowest rate of citations issued, while Hispanics have the highest rate of citations issued. In Colorado, all rates are similar with little noticeable discrepancy.

## Warnings

```{sql}
#| connection: con_traffic
#| output.var: "warnings_table"
#| code-fold: true
#| code-summary: "Warnings Table"

-- Maryland warnings 
SELECT 'Maryland' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(warning_issued) AS num_warnings,
  SUM(warning_issued) * 1.0 / COUNT(*) AS rate_of_warnings -- calculate rate of warnings
FROM md_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 

UNION

-- Arizona warnings
SELECT 'Arizona' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(warning_issued) AS num_warnings,
  SUM(warning_issued) * 1.0 / COUNT(*) AS rate_of_warnings -- calculate rate of warnings
FROM az_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 

UNION 

-- Colorado warnings
SELECT 'Colorado' AS state, 
  subject_race,
  COUNT(*) AS total_stops,
  SUM(warning_issued) AS num_warnings,
  SUM(warning_issued) * 1.0 / COUNT(*) AS rate_of_warnings -- calculate rate of warnings
FROM co_statewide_2020_04_01
WHERE subject_race NOT IN ('other', 'unknown') -- remove other and unknown
GROUP BY subject_race 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
head(warnings_table)
```

The SQL query creates a table with information regarding the rate of warnings issued from traffic stops in Maryland, Colorado, and Arizona, categorized by race. The information is filtered to remove data that had race represented as 'other' or 'unknown'. It is compiled into one table, where it displays each state, subject_rate, and rate_of_warnings. The data from these columns is then used to create a visualization.

```{r warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: "Plot of warnings"
#| fig-alt: "The bar graph represents the rate of warnings given across the races Black, Hispanic, Asian/Pacific Islander, and White in the states Maryland, Arizona, and Colorado. Maryland issues the highest rate of warnings while Colorado issues the least across all races." 
ggplot(warnings_table, aes(x = subject_race, y = rate_of_warnings, fill = state)) +
  geom_col(position = "dodge") + # create bar plot using rate of warnings and race
  labs( # plot labels
    title = "Rate of warnings given by race across the states \n
    Arizona, Colorado, and Maryland",
    x = "State",
    y = "Rate of Warnings issued"
  )
```

Unlike arrests and citations, Maryland has the highest rate of warnings issued across all races. Asian/Pacific Islanders have the highest rate of warnings, while Hispanics have the lowest rate. In Arizona, White individuals have the highest rate of warnings while Hispanics have the lowest, just slightly lower than Asian/Pacific Islander. In Colorado, numbers are very similar once again, except for a slight bump in warnings given to White people.

## Conclusion

In order to create the visualizations of arrests made, citations issued, and warnings given by race across the states of Arizona, Maryland, and Colorado in traffic stops, I utilized SQL queries. Each dataset had a variable pertaining to arrests, citations, or warnings. I extracted the number of each punishment given, grouping by race then dividing by the number of total stops in order to calculate the rate of arrests, warnings, and citations given. I then used union to combine data and create tables for each type of punishment given. I also removed data points that labelled race as 'other' or 'unknown', as this data was not helpful to the analysis we were completing. An issue that makes the findings difficult to interpret is that the data did not include race breakdown across the three states that I chose to use. This would have made for even clearer data with more meaning, because it would be the rate of punishment from traffic stops, with regard to the percent of the population that the race made up.

## References

“Data.” The Stanford Open Policing Project, 2020, https://openpolicing.stanford.edu/data/. Accessed 8 May 2025.

Pierson, Emma, Camelia Simoiu, Jan Overgoor, Sam Corbett-Davies, Daniel Jenson, Amy Shoemaker, Vignesh Ramachandran, et al. 2020. “A Large-Scale Analysis of Racial Disparities in Police Stops Across the United States.” *Nature Human Behaviour*, 1–10.
