---
title: "Bobs Burgers"
description: |
  Analysis of Bob's Burgers dialogue data from TidyTuesday
author: Connor Wang
date: February 15, 2025
format: html
---

```{r echo=FALSE, results = "hide", message=FALSE, warning=FALSE}
library(tidyverse)

episode_metrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-19/episode_metrics.csv')

episode_metrics
```

```{r echo=FALSE}
ggplot(episode_metrics, aes(x = question_ratio, y = exclamation_ratio)) +
  geom_point(color = "blue") + 
  labs(
    title = "Question Ratio vs. Exclamation Ratio in Bob's Burgers episodes",
    x = "Ratio of lines with questions",
    y = "Ratio of lines with exclamations"
  )
```

Based on the scatterplot above, there is no clear correlation between amount of questions and exclamations.

## References

Poncest. “Poncest/Bobsburgersr: Collection of Datasets from the Bob’s Burgers Animated Sitcom.” GitHub, github.com/poncest/bobsburgersR. Accessed 15 Feb. 2025.

Rfordatascience. “Tidytuesday/Data/2024/2024-11-19/Readme.Md at Main · Rfordatascience/Tidytuesday.” GitHub, github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-11-19/readme.md. Accessed 15 Feb. 2025.
